Red Wine Quality Exploration by Taghreed
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages 

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

setwd("F:/New/Data Analyst/term2-project2")
library(ggplot2)
library(dplyr)
library(RColorBrewer)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
red <- read.csv('wineQualityReds.csv')
```

```{r echo=FALSE, Dimensions}
dim(red)
```

```{r echo=FALSE, Structure}
str(red)
```

```{r echo=FALSE, Summary}
summary(red)
```

> Our dataset consists of 13 variables, with 1599 observations.

```{r echo=FALSE, Quality_Ordered}
red$quality <- ordered(red$quality)
str(red)
```

# Univariate Plots Section

```{r echo=FALSE, Univariate_Plot_Quality}
ggplot(aes(x = quality), data = red) +
  geom_bar()
```

```{r echo=FALSE}
table(red$quality)
```

In this dataset there is no wine has quality > 8 or < 3.
Most of wine have quality of 5 or 6.

```{r echo=FALSE, Univariate_Plot_pH}
ggplot(aes(x = pH), data = red) +
  geom_histogram()
```

Most of the wine have pH between 3.2 - 3.5.

```{r echo=FALSE, Univariate_Plot_Alcohol}
ggplot(aes(x = alcohol), data = red) +
  geom_histogram()
```

The Distribution of alcohol is Right-Skewed.
Most of the wine have alcohol between 9 - 12.

```{r echo=FALSE, Univariate_Plot_sulphates}
ggplot(aes(x = sulphates), data = red) +
  geom_histogram() +
  scale_x_log10()
```

The Distribution of sulphates was Right-Skewed with long tail, So used the log_scale.
Most of the wine have sulphates between 0.5 - 0.8.

```{r echo=FALSE, Univariate_Plot_density}
ggplot(aes(x = density), data = red) +
  geom_histogram()
```

Most of the wine have density between 0.995 - 1.

```{r echo=FALSE, Univariate_Plot_chlorides}
ggplot(aes(x = chlorides), data = red) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 0.2))
```

The Distribution of chlorides had right long tail, but focusing on the bulk of the data between (0, 0.2), the distribution is normal
Most of the wine have chlorides between 0.05 - 0.1.

```{r echo=FALSE, Univariate_Plot_residual_sugar}
ggplot(aes(x = residual.sugar), data = red) +
  geom_histogram(bins = 40) +
  scale_x_continuous(limits = c(0, 4))
```

```{r echo=FALSE, Summary_residual_sugar}
summary(red$residual.sugar)
```

The Distribution of residual sugar had right long tail, but focusing on the bulk of the data between (0, 4), the distribution seems normal with a peak around 2 
Most of the wine have residual sugar between 1.9 - 2.6.

```{r echo=FALSE, Univariate_Plot_total_sulfur}
ggplot(aes(x = total.sulfur.dioxide), data = red) +
  geom_histogram() +
  scale_x_log10()
```

The Distribution of total sulfur dioxide was Right-Skewed & there was outliers between 250 - 300.
The Distribution is normal in log scale.

```{r echo=FALSE, Univariate_Plot_volatile_acidity}
summary(red$volatile.acidity)

ggplot(aes(x = volatile.acidity), data = red) +
  geom_histogram() +
  scale_x_continuous(limits = c(0,1))
```

The Distribution of volatile acidity was Right-Skewed & there was outliers.
But focusing on the bulk of the data between (0, 1), the distribution is normal with 3 peaks in range(0.3 - 0.65)

```{r echo=FALSE, Univariate_Plot_fixed_acidity}
ggplot(aes(x = fixed.acidity), data = red) +
  geom_histogram() +
  scale_x_log10()
```

```{r echo=FALSE, Univariate_Plot_citric_acid}
summary(red$citric.acid)

ggplot(aes(x = citric.acid), data = red) +
  geom_histogram()

dim(red[red$citric.acid == 0,])[1]
```

The Distribution of citric acid was Right-Skewed with ups & downs.
There are 132 observation with zero citric acid.

# Univariate Analysis

### What is the structure of your dataset?
Our dataset consists of 13 variables, with 1599 observations.

Input variables (based on physicochemical tests):
   1 - fixed acidity (tartaric acid - g / dm^3)
   2 - volatile acidity (acetic acid - g / dm^3)
   3 - citric acid (g / dm^3)
   4 - residual sugar (g / dm^3)
   5 - chlorides (sodium chloride - g / dm^3
   6 - free sulfur dioxide (mg / dm^3)
   7 - total sulfur dioxide (mg / dm^3)
   8 - density (g / cm^3)
   9 - pH
   10 - sulphates (potassium sulphate - g / dm3)
   11 - alcohol (% by volume)
   
Output variable (based on sensory data): 
   12 - quality (score between 0 and 10)

### What is/are the main feature(s) of interest in your dataset?
The output variable (Quality) 

### What other features in the dataset do you think will help support your \ investigation into your feature(s) of interest?
Other input variables and how they affect the quality of wine.

### Did you create any new variables from existing variables in the dataset?
No

### Of the features you investigated, were there any unusual distributions? \ Did you perform any operations on the data to tidy, adjust, or change the form \ of the data? If so, why did you do this?
Some of distribution are right-skewed with long tail, So i try to use log scaling or focus on the bulk of the data to check the distribution.

# Bivariate Plots Section

```{r echo=FALSE, Bivariate_Plots}
library('GGally')
ggpairs(red[, c(2:13)])
```

check the correlation between variables.

```{r echo=FALSE, correlation}
# round(cor(red[,2:12]), 3)
ggcorr(red[,2:12], 
       label = TRUE, 
       label_alpha = TRUE)
```

```{r echo=FALSE, Bivariate_Plot_quality_alcohol}
ggplot(aes(x = quality, y = alcohol), data = red) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

by(red$alcohol, red$quality, summary)
```

As alcohol increases, the quality increases.

```{r echo=FALSE, Bivariate_Plot_quality_volatile.acidity}
ggplot(aes(x = quality, y = volatile.acidity), data = red) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

by(red$volatile.acidity, red$quality, summary)
```

As volatile.acidity decreases, the quality increases.

```{r echo=FALSE, Bivariate_Plot_quality_volatile.acidity}
ggplot(aes(x = quality, y = citric.acid), data = red) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

by(red$citric.acid, red$quality, summary)
```

As citric.acid increases, the quality increases.

```{r echo=FALSE, Bivariate_Plot_quality_volatile.acidity}
ggplot(aes(x = quality, y = sulphates), data = red) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)
```

As sulphates increases, the quality increases.

```{r echo=FALSE, Bivariate_Plot_Alcohol_density}
ggplot(aes(x = alcohol, y = density), data = red) +
  geom_point(fill = I('#F79420'), color = I('black'), shape = 21, alpha = 1/5) +
  stat_smooth(method = 'lm')
```

As alcohol increases, the density decreases.

```{r echo=FALSE, Bivariate_Plot_pH_citric.acid}
ggplot(aes(x = citric.acid, y = pH), data = red) +
  geom_point(fill = I('#099DD9'), color = I('black'), shape = 21, alpha = 1/5) +
  geom_smooth(method = 'lm', color = I('orange'))
```

As citric acid increases, the pH decreases.

```{r echo=FALSE, Bivariate_Plot_volatile.acidity_citric.acid}
ggplot(aes(x = citric.acid, y = volatile.acidity), data = red) +
  geom_point(fill = I('#099DD9'), color = I('black'), shape = 21, alpha = 1/5) +
  geom_smooth(method = 'lm', color = I('orange'))
```

As citric acid increases, the volatile.acidity decreases.

```{r echo=FALSE, Bivariate_Plot_fixed.acidity_pH}
ggplot(aes(x = fixed.acidity, y = pH), data = red) +
  geom_point(fill = I('#099DD9'), color = I('black'), shape = 21, alpha = 1/5) +
  geom_smooth(method = 'lm', color = I('orange'))
```

As fixed.acidity increases, the pH decreases.

```{r echo=FALSE, Bivariate_Plot_residual.sugar_density}
ggplot(aes(x = residual.sugar, y = density), data = red) +
  geom_point(fill = I('#099DD9'), color = I('black'), shape = 21, alpha = 1/5) +
  geom_smooth(method = 'lm', color = I('orange')) +
  scale_x_continuous(limits = c(0, 5))
```

As residual.sugar increases, the density increases and this is clear in range (0-5) of residual.sugar.

```{r echo=FALSE, Bivariate_Plot_fixed.acidity_citric.acid}
ggplot(aes(x = citric.acid, y = fixed.acidity), data = red) +
  geom_point(fill = I('#099DD9'), color = I('black'), shape = 21, alpha = 1/5) +
  geom_smooth(method = 'lm', color = I('orange'))
```

As citric.acid increases, the fixed.acidity increases.

```{r echo=FALSE, Bivariate_Plot_fixed.acidity_density}
ggplot(aes(x = fixed.acidity, y = density), data = red) +
  geom_point(fill = I('#099DD9'), color = I('black'), shape = 21, alpha = 1/5) +
  geom_smooth(method = 'lm', color = I('orange'))
```

As fixed.acidity increases, the density increases.

```{r echo=FALSE, Bivariate_Plot_free.sulfur.dioxide_total.sulfur.dioxide}
ggplot(aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide), data = red) +
  geom_point(fill = I('#099DD9'), color = I('black'), shape = 21, alpha = 1/5) +
  geom_smooth(method = 'lm', color = I('orange')) +
  scale_x_continuous(limits = c(0, 40)) +
  scale_y_continuous(limits = c(0, 200))
```

As free.sulfur.dioxide increases, the total.sulfur.dioxide increases.


# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \ investigation. How did the feature(s) of interest vary with other features in \ the dataset?

The quality is postively correlated with alcohol, citric.acid & sulphates.
The quality is negatively correlated with volatile.acidity.

### Did you observe any interesting relationships between the other features \ (not the main feature(s) of interest)?

According correlation Matrix

|   Variables                              | Correlation |
|------------------------------------------|-------------|
|pH ~ fixed.acidity                        | -0.7        |
|volatile.acidity ~ citric.acid            | -0.6        |
|alcohol ~ density                         | -0.5        |
|pH ~ citric.acid                          | -0.5        |
|citric.acid ~ fixed.acidity               |  0.7        |
|density ~ fixed.acidity                   |  0.7        |
|free.sulfur.dioxide ~ total.sulfur.dioxide|  0.7        |
|residual.sugar ~ density                  |  0.4        |

Most of these relationships can be represented logically 
- More acidity leeds to low pH
- More alcohol less dense
- More sugar more dense

### What was the strongest relationship you found?
According correlations between variables the strongest relationship is between pH ~ fixed.acidity (negatively).
Also (citric.acid ~ fixed.acidity) & (density ~ fixed.acidity) are strong.

Related to the output quality, volatile.acidity is negatively correlated with quality. 

# Multivariate Plots Section

Consider wine with quality(7-8) is high quality wine & quality(5-6) which represnt the most of the data is moderate quality wine.

```{r echo=FALSE, Multivariate_Plot_citric.acid_volatile.acidity}
ggplot(aes(x = citric.acid, y = volatile.acidity), data = red) +
  geom_point(aes(color = quality), alpha = 1/2) +
  scale_x_continuous(limits = c(0, 0.8)) + 
  scale_colour_brewer(palette = "Dark2",
    guide = guide_legend(title = 'Quality', reverse = T,
    override.aes = list(alpha = 1, size = 2)))
```

Most of High quality wine seem to have high citric.acid & low volatile.acidity.

```{r echo=FALSE, Multivariate_Plot_citric.acid_fixed.acidity}
ggplot(aes(x = citric.acid, y = fixed.acidity), data = red) +
  geom_point(aes(color = quality), alpha = 1/2) + 
  scale_colour_brewer(palette = "Dark2",
    guide = guide_legend(title = 'Quality', reverse = T,
    override.aes = list(alpha = 1, size = 2)))
```

High quality wine seem to have high citric.acid & high fixed.acidity.

```{r echo=FALSE, Multivariate_Plot_fixed.acidity_density}
ggplot(aes(x = fixed.acidity, y = density), data = red) +
  geom_point(aes(color = quality), alpha = 1/2) + 
  scale_colour_brewer(palette = "Dark2",
    guide = guide_legend(title = 'Quality', reverse = T,
    override.aes = list(alpha = 1, size = 2)))
```

Most of High quality wine seem to have lower density compared to moderate wine quality.

```{r echo=FALSE, Multivariate_Plot_citric.acid_pH}
ggplot(aes(x = citric.acid, y = pH), data = red) +
  geom_point(aes(color = quality), alpha = 1/2) + 
  scale_colour_brewer(palette = "Dark2",
    guide = guide_legend(title = 'Quality', reverse = T,
    override.aes = list(alpha = 1, size = 2)))
```

High quality wine seem to have high citric.acid & low pH

```{r echo=FALSE, Multivariate_Plot_residual.sugar_density}
ggplot(aes(x = residual.sugar, y = density), data = red) +
  geom_point(aes(color = quality), alpha = 1/2) +
  scale_x_continuous(limits = c(1, 5)) + 
  scale_colour_brewer(palette = "Dark2",
    guide = guide_legend(title = 'Quality', reverse = T,
    override.aes = list(alpha = 1, size = 2)))
```

High quality wine seem to have low residual.sugar & low density.

```{r Building the Linear Model}
library(memisc)

m1 <- lm(as.numeric(quality) ~ alcohol, data = red)
m2 <- update(m1, ~ . + volatile.acidity)
m3 <- update(m2, ~ . + citric.acid)
m4 <- update(m3, ~ . + sulphates)

mtable(m1, m2, m3, m4)
```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \ investigation. Were there features that strengthened each other in terms of \ looking at your feature(s) of interest?


### Were there any interesting or surprising interactions between features?


### OPTIONAL: Did you create any models with your dataset? Discuss the \ strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
library(scales)

ggplot(aes(x = quality, fill = quality), data = red) +
  geom_bar(aes(y = (..count..)/sum(..count..))) +
  scale_y_continuous(labels = percent) +
  scale_fill_brewer(type = "seq", palette = "Blues",
    guide = guide_legend(title = 'Quality', reverse = T)) +
  theme_dark() +
  ggtitle('Percent of Wine Quality') +
  xlab('Quality') +
  ylab('Percent (%)')
```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}
ggplot(aes(x = quality, y = volatile.acidity, fill = quality), data = red) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4) + 
  scale_fill_brewer(palette = "Blues") +
  guides(fill=FALSE) + 
  ggtitle('') +
  xlab('Quality') +
  ylab(expression(Volatile~Acidity~(g/dm^{3})))
```

```{r echo=FALSE,}
ggplot(aes(x = quality, y = citric.acid, fill = quality), data = red) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4) + 
  scale_fill_brewer(palette = "Blues") +
  guides(fill=FALSE) +
  ggtitle('') +
  xlab('Quality') +
  ylab(expression(Citric~Acid~(g/dm^{3})))
```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}
ggplot(aes(x = citric.acid, y = volatile.acidity), data = red) +
  geom_point(aes(color = alcohol), alpha = 1/2) +
  scale_x_continuous(limits = c(0, 0.8)) + 
  scale_colour_gradient(low = "blue", high = "red", guide = "colourbar",
  aesthetics = "colour", guide_legend(title = 'Alcohol')) +
  ggtitle('Volatile Acidity vs Citric Acid vs Alcohol') +
  xlab(expression(Citric~Acid~(g/dm^{3}))) +
  ylab(expression(Volatile~Acidity~(g/dm^{3})))
```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!